\chapter{Preliminaries}

\section{Set Functions}

% co je set funkce
Set functions are very general objects studied extensively in countless fields, including
matroid theory, game theory utility functions and countless others. % TODO: citace a konkrétní příklady


% definice set funkce
\begin{defi}[Set function]
	Let $ N $ be a finite set.
	A \emph{set function} on the set $ N $ is any function $ f: \pot N \to \R $.
\end{defi}

Though the definition of a set function consideres a general finite set $ N $, in this thesis,
we will for simplicity only consider set functions on the set $ N = \{1, 2, \ldots, n\} $.
This simplification is common in the literature,  % TODO: citace; jinak, proste jsou izomorfni
as it allows us to be more concrete when discussing set functions, while
the results we show can be easily generalized to the case of a general finite set $ N $.

% k čemu jsou
For example, $ N $ can represent a set of \emph{goods}, while the set function $ f $ can represet a \emph{utility} an agent receives from obtaining a subset of the goods.

% monotonicita
When studying set functions modelling real-world problems, we can often assume that the set function has certain properties.
One of the most natural and commonly observed of such properties is \emph{monotonicity}.

\begin{defi}[Monotonicity]
	A set function $ f: \pot N \to \R $ is \emph{monotonically non-decreasing} if \[
		\left( \forall S \subseteq T \subseteq N \right)\quad \fce{f}{S} \leq \fce{f}{T},
	\]
	and \emph{monotonically non-increasing} if \[
		\left( \forall S \subseteq T \subseteq N \right)\quad \fce{f}{S} \geq \fce{f}{T}.
	\]
\end{defi}

% ukázat na utilitě
In our example of a set function representing utility, assuming the utility of each good by itself is non-negative, the monotonicity property is equivalent to the assumption that acquiring more goods can never decrease the utility of an agent.
% TODO: ne vždycky

% superadditivita
Another very common and natural property of many set functions is \emph{superadditivity}.

% TODO: jak to udělat s tím že máme i subadditivitu? -- nejdriv super, pak note: sub
\begin{defi}[Superadditivity]
	A set function $ f: \pot N \to \R $ is \emph{superadditive} if \[
		\left( \forall S, T \subseteq N; S \cap T = \emptyset \right)\quad \fce{f}{S \cup T} \geq \fce{f}{S} + \fce{f}{T}.
	\]
\end{defi}

% ukázat na utilitě
Getting back to our example with utility, superadditivity means that the utility of obtaining two sets of goods together is at least as high as the sum of the utilities of obtaining each of the sets separately.
In other words, a set is at least as valuable as the sum of its parts.

% supermodularita
\begin{defi}[Supermodularity]
	A set function $ f: \pot N \to \R $ is \emph{supermodular} if \[
		\left( \forall S, T \subseteq N \right)\quad \fce{f}{S \cup T} + \fce{f}{S \cap T} \geq \fce{f}{S} + \fce{f}{T}.
	\]
\end{defi}
% TODO: proč je supermodularita sexy
% TODO: hrozně sexy struktura

% submodularita
\begin{defi}[Submodularity]
	A set function $ f: \pot N \to \R $ is \emph{submodular} if \[
		\left( \forall S, T \subseteq N \right)\quad \fce{f}{S \cup T} + \fce{f}{S \cap T} \leq \fce{f}{S} + \fce{f}{T}.
	\]
\end{defi}
% TODO: vztah se supermodularitou
% TODO: proč submodularita je taky sexy
% TODO: greedy algoritmy a matroidy

% třídy funkcí
When discussing set functions with certain properties, it is often useful to consider \emph{classes} of set functions with those properties.
We will denote $ \cAdditive, \cSuperadditive $ and $ \cSupermodular $ the classes of all set functions on the set $ N $ that are additive, superadditive, and supermodular (sometimes also called \emph{convex}), respectively.

% TODO: capitalization?
\section{Cooperative Game Theory}

% co je to kooperativní teorie her
Cooperative game theory is a field of mathematics that studies situations in which agents can form cooperating groups, called \emph{coalitions}.
These coalitions receive a \emph{payoff} from the game, which is then somehow divided among the agents in the coalition.
The goal of cooperative game theory is to study the \emph{stability} of the formed coalitions, and the \emph{distribution} of the payoff among the agents.

% definice kooperativní hry
\begin{defi}[Cooperative game]
	A \emph{cooperative game} is a pair $ \left( N,v \right) $, where $ N $ is a finite set of \emph{players} and $ v: \pot N \to \R $ is the \emph{characteristic function} of the game.
	Further, we require that $ \fce{v}{\emptyset} = 0 $.
\end{defi}

When discussing coalitions, we usually use the terms \emph{grand coalition} for the coalition of all players $ N $, and \emph{singletons} for coalitions containing a single player.

% TODO: vztah se set functs
From the definition it is clear that the characteristic function is a set function on the set $ N $, with the additional requirement that the value of the empty coalition is zero.

% supermodularita => superadditivita
% TODO: kopirovat do setfunkcí
\begin{prop}[ ]
	Let $ \left( N,v \right) $ be a cooperative game.
	If the characteristic function $ v $ is supermodular, then it is superadditive.
\end{prop}

% TODO: nebudeme chtít tohle mít pro všechny set funkce? chceme to řešit v preliminaries nebo radši až potom v definici lower/upper?
\begin{proof}
	Suppose $ S,T \subseteq N $, such that $ S \cap T = \emptyset $.
	We will use the definition of supermodularity, along with the fact that $ \fce{v}{\emptyset} = 0 $.
	We get \[
		\fce{v}{S} + \fce{v}{T} \leq \fce{v}{S \cup T} + \fce{v}{S \cap T} = \fce{v}{S \cup T} + \fce{v}{\emptyset} = \fce{v}{S \cup T},
	\]
	which is exactly the definition of superadditivity.
\end{proof}

% payoff vector
In cooperative game theory, we usually assume that the grand coalition has formed, i.e. all the players are cooperating, then we discuss how the value of the grand coalition, $ \fce{v}{N} $, shall be split between the players, and then investigate whether the formed coalition is stable, or not.
The payoff of players is described by a payoff vector.

\begin{defi}[Payoff vector]
	Let $ \left( N,v \right) $ be a cooperative game.
	A \emph{payoff vector} is $ x \in \R^n $.
	The payoff vector is \emph{efficient} if \[
		\sum_{i = 1}^{n} x_i = \fce{v}{N}.
	\]
\end{defi}

For simplicity, we usually shorten the notation of sum over the values of the payoff vector.
For a general coalition $ S \subseteq N $, we denote \[
	\fce{x}{S} \deq \sum_{i \in S}^{} x_i. % TODO: tohle lip: pouzivame rozsireni setfunkci bla bla; also, dat vic na zacatek?
\]
% efficiency
In this thesis, we will only consider efficient payoff vectors, as we want to distribute the entire value of the grand coalitions.
Other notions of distribution where the payoff vector needn't be efficient is sometimes studied, but we will not discuss them here. % TODO: citace

% solution concept
We are interested in studying payoff vectors with certain properties, such as fairness, stability, etc.
For this, we define solution concepts as the sets of payoff vectors with given properties.
Formally

\begin{defi}[Solution concept]
	A \emph{solution concept} is a function $ f: \games n \to \pot{\R^n} $.
\end{defi}

% TODO: masirovat
Some solution concepts, so-called \emph{multi-point solution concepts}, give sets of payoff vectors of varying size, depending on the game.
However, some important solution concepts are so-called \emph{single-point}, that is, they map each game to exactly one vector.
In those cases, we simplify the notation to leave out the set, yielding a solution concept with the signature $ f: \games n \to \R^n $.

The first important solution concept we will mention here is the \emph{core}.
The core the payoff vectors, which encourage stability of the grand coalition.
In other words, when distributing $ \fce{v}{N} $ according to a payoff vector from the core, the players don't have incentive to leave the grand coalition.

\begin{defi}[Core]
	The \emph{core} is a multi-point solution concept defined as \[
		\core v \deq \left\{ x \in \R^n \suchthat \fce{x}{N} =\fce{v}{N} \land \left( \forall S \subseteq N \right)\! \left( \fce{x}{S} \geq \fce{v}{S} \right) \right\}.
	\]
\end{defi}

If the core of a game is empty, it tells us that no matter how we distribute the value of $ \fce{v}{N} $, there will always be a set of players that would be better off leaving the grand coalition, so it is unstable, and thus (assuming perfect rationality of the players), it won't form.

% shapley
However, the core doesn't consider the fairness of the distribution.
For that, we would like another solution concept, that would divide the value of the grand coalition proportionally to the contributions of each player.
By contributions, we mean the extra value the coalition generates, when a player joins. Formally, for $ i \in N $ and $ S \subseteq N \setminus \left\{ i \right\} $, the contribution of $ i $ to $ S $ is $ \fce{v}{S \cup \left\{ i \right\}} - \fce{v}{S} $.

\begin{defi}[Shapley value]
	The \emph{Shapley value} is a single-point solution concept defined as \[
		\shapley v_i \deq \sum_{S \subseteq N \setminus \left\{ i \right\}}^{} \frac{s! \left( n - s - 1 \right)!}{n!} \left( \fce{v}{S \cup \left\{ i \right\}} - \fce{v}{S} \right),
	\]
	where $ s \deq \absolute{S} $.
\end{defi}

So, is the Shapley value fair?
That obviously depends on your definition of fairness, which can be very subjective.
However, the Shapley value is generally accepted as fair, as it satisfies many natural qualities that a reasonable fair solution concept should satisfy.

% TODO: citace
\begin{prop}[Characterization of the Shapley value]
	The Shapley value $ \Shapley $ is the unique function satisfying the following properties: \begin{enumerate}[ ]
		\item efficience, $ \sum_{i \in N}^{} \fce{f_i}{v} = \fce{v}{N} $,
		\item symmetry, \[
				\left( \forall i,j \in N \right)\! \left( \left( \forall S \subseteq N \setminus \left\{ i,j \right\} \right)\! \left( \fce{v}{S \cup i} = \fce{v}{S \cup j} \right) \implies \fce{f_i}{v} = \fce{f_j}{v} \right),
			\]
		\item null player, $ \left( \forall i \in N \right)\! \left( \left( \forall S \subseteq N \setminus i \right)\! \left( \fce{v}{S} = \fce{v}{S \cup i} \right) \implies \fce{f_i}{v} = 0 \right)   $,
		\item additivity, $ \left( \forall v,w \in \games n \right)\! \left( \fce{f}{v + w} = \fce{f}{v} + \fce{f}{w} \right) $.
	\end{enumerate}
\end{prop}
% TODO: další vlastnosti SV?

\section{Reinforcement Learning\filip{20.4}}
\label{sec:rl}

% RL je způsob jak udělat agenta
Reinforcement learning (RL) is a technique for designing \emph{agents} which learn to perform tasks.
The agent interacts with an environment, and receives feedback in the form of rewards.
The goal of the agent is to learn a policy that maximizes the expected future reward.
\emph{Deep} reinforcement learning extends this concept by incorporating deep neural networks to the agent.

% my budeme dělat jen trochu RL
RL is an extensive field, which shall be discussed in this thesis only to the extent required for our purposes.
Similarly, we do not further discuss the details of training deep neural networks.
The reader is encouraged to look deeper into the literature for more information, for example \cite{sutton2018reinforcement,bengio2017deep}.

% % co to je
% Machine learning refers to a field of Computer Science, where we study algorithms that aren't explicitly designed to perform a specific task, but rather are programmed to learn the expected behavior by analyzing data.
% Online learning is a subfield of machine learning, where an \emph{agent} interacts with a black-box \emph{environment}, and receives \emph{feedback} on its chosen actions.
% Online learning is a general model, which can be used to model many different situations, including those with multiple agents.
% However, in this thesis, we focus on the online learning of a single agent.

% agent+environment
% TODO: používám tady "game", což by chtělo někde nahoře zmínit co je, nebo to změnit na jiné slovo
Let us first formally define the model used when discussing RL.
The environment is a black box to the agent, which holds a \emph{state}.\david{tahle věta se mi moc nelíbí jako první hned za "teď pojdme dělat věci formálně", ale moc nevim jak jí změnit. Možná něco o tom, že si env drží interně stav?}\filip{Uhh o tom ta věta je ne? :D}\david{jako ano, akorát se mi nelíbí ta formulace. Prostě za formally by měla být pauza, než následuje slovo some :D}\filip{ok, akorát "internal" mi zavání POMDP...}\david{Takhle dobrý, možná by se dalo říct which is in some state, ale tohle taky fajn.}\filip{Davide, takhle to ale bylo před tím, to jsem přesně měnil ne? Přesně jsi říkal že mezi "formally" a "some" má být hodně pauza. Lol :D}\david{:D :D :D :D :D Ajo :D No tak zjevně jsme full circle, udělej to jak myslíš a já to budu dál hatit xD}
The agent interacts with the environment via two functions -- the \emph{transition function} $ \tran $, which informs the agent what state the envrionment moves to after he chooses an action,
and the \emph{reward function} $ \rew $, which tells him the \emph{reward} he received for choosing an action (the reward also possibly depends on the \emph{outcome} of the action, i.e. the new state).
The goal of the agent is to maximize the cumulative reward received over time.
We will formalize this as a \emph{Markov decision process}.
Note that in the definition of the Markov decision process, we further require a \emph{discount factor} $ \gamma $, which will be discussed later.

% MDP?
\begin{defi}[Markov decision process]\label{def:mdp}
	A \emph{Markov decision process} (MPD) is a tuple $ \left( \actions, \states, \tran, \rew, \gamma \right) $, where \begin{enumerate}[ ]
		\item $ \actions $ is a non-empty set of \emph{actions},
		\item $ \states $ is a non-empty set of \emph{states},
		\item $ \tran: \states \times \actions \to \dist \states $ is the \emph{transition function},\filip{Jinak jak jsme se bavili o tom že straka má $ p(s_{t+1},r_{t+1} | s_t) $, tak to je sice pravda, a není to úplně ekvivalentní, ale moje definice je zase používaná třeba na wikipedii, takže to asi prostě je jedno :D}\david{Jo tohle je podle mě super}
		\item $ \rew: \states \times \actions \times \states \to \R $ is the \emph{reward function},
		\item $ \gamma \in \left[ 0,1 \right] $ is the \emph{discount factor}.
	\end{enumerate}
\end{defi}
% TODO: daviduv oblibeny obrazek

At each step $ t $, the agent chooses an action $ a_t \in \actions $, is informed of the new state $ s_{t+1} \sim \fce{\tran}{s_{t}, a_t} $ of the environment, and receives reward $ \rew_t = \fce{\rew}{s_{t}, a_t, s_{t+1}} $.
Notice that the reward depends not only on the initial state and the chosen action, but also on the final state.
To formalize the agent, we define a \emph{policy}, which is a function that gives a \emph{distribution on the next actions} to take, depending on the current state.

\begin{defi}[Policy]
	\label{def:policy}
	A \emph{policy} of an agent in an environment $ \left( \actions, \states, \tran, \rew, \gamma \right) $ is a function $ \pi: \states \to \dist \actions $.
\end{defi}

% TODO: \filip{Co dát aspoň část tohohle odstavce až pod def returnu?}
% \david{jakože klidně, přijde mi to takhle dobrý. Já rád omáčku před definicí}
As the agent plays, sampling his actions according to the policy $ \pi $, he receives some reward at every step.
We are interested in quantifying \emph{how well} a policy performs over time, which translates to \emph{how much reward} the agent receives \emph{overall}.
\david{emph spíš na overall?}\filip{já to tam dával aby to bylo symetrické k tomu \emph{how well}. Takhle už je to asi moc žejo?}\david{ok, takhle taky dobrý. Ale prostě overall je to důležitý}
However, formally, the agent plays indefinitely.
To make the total expected cumulative reward finite, we discount the future rewards exponentially, by a factor of $ \gamma^t $.\footnote{This gives a finite sum, assuming $ \gamma < 1 $ and the rewards are bounded.}
This is similar to human behavior, often prefering smaller immediate rewards over larger rewards far in the future.
To capture environments with a finite horizon of $ T $ steps, we set the rewards for times $ t > T $ to zero.

% TODO: vysvětlivka co znamená \E_\pi
\begin{defi}[Return]
	Let $ \left( \actions, \states, \tran, \rew, \gamma \right) $ be a MDP.
	The agent chooses actions $ \left( a_t \right)_{t=0}^\infty $ according to a policy $ \pi $, reaching states $ \left( s_t \right)_{t=0}^\infty $ and receiving rewards $ \left( \rew_t \right)_{t=0}^\infty $.
	Then the \emph{return} of the agent from time $ \tau $ is \[
		\return_T \deq \sum_{t=\tau}^{\infty} \gamma^{t-\tau} \rew_t.
	\]
\end{defi}

Since the transition function and the policy are stochastic, the states, actions, and rewards, and thus the return, are random variables.
When constructing a policy, our goal will be to maximize $ \fceb{\E}{\return_0} $, where the expectation is taken over the transition function and the policy in each step.

The discount factor $ \gamma $ decreases the value of future rewards, which serves two purposes.
It ensures that the return is finite, assuming the rewards are bounded, since $ \sum_i \gamma^i $ converges for $ \gamma \in \left( 0,1 \right) $.
Further, it makes the agent prefer immediate rewards over future rewards, which sometimes helps the training algorithms to converge faster, as is discussed below.

We further want to quantify the performance of a policy in a given state.

\begin{defi}[State value]\label{def:state-value}
	Suppose the agent takes steps in an environment $ \left( \actions, \states, \tran, \rew, \gamma \right) $ according to a policy $ \pi $.
	The \emph{value} of a state $ s $ is \[
		\fce {\stval} {s} \deq \E_\pi \left[ \return_0 \suchthat s_0 = s \right].
	\]
\end{defi}

Note that since the process is Markovian, the value of a state $ s $ is equal to $ \E_\pi \left[ \return_t \suchthat s_t = s \right] $, for any $ t \in \N $.
Finally, let us define the value of choosing an action in a state.

\begin{defi}[Action-state value]
	Suppose the agent takes steps in an environment $ \left( \actions, \states, \tran, \rew, \gamma \right) $ according to a policy $ \pi $.
	The \emph{action-state value} of a state $ s $ and an action $ a $ is \[
		\fce{q_\pi}{s, a} \deq \E_\pi \left[ \return_0 \suchthat s_0 = s, a_0 = a \right].
	\]
\end{defi}

This quantity is also referred to as the \emph{q-value} of an action in a state.
Action-state values approximate the change in the state value, when a given action in a given state is assigned a larger probability.
This is crucial in learning, when deciding how to alter a policy to maximize the expected return.

% TODO: tohle PPO
\subsection{PPO\filip{20.4}}

% existuje spoustu způsobů, my vybíráme PPO + NN
There are many reinforcement learning algorithms, each with its own strengths and weaknesses.
In our experiments, we use the \emph{Proximal Policy Optimization} (PPO) algorithm. % TODO: citace
PPO is a state-of-the-art algorithm, which has been empirically shown to perform well in various environments. % TODO: citacee
Importantly for our purposes, PPO is able to work with continuous state spaces, which is required in our application.%
\footnote{This is not specific to PPO, but rather to the fact that we are using neural networks.}

% PPO funguje actor-critic
Let us now delve into the specifics of the PPO algorithm.
PPO belongs to a family of so-called actor-critic algorithms.
% PPO uses an actor-critic architecture.
It uses two deep neural networks, the \emph{actor} and the \emph{critic}.
% The \emph{actor} and the \emph{critic} are used to approximate the policy and state value functions, respectively.
The actor tries to learn a policy $ \pi $ which maximizes the expected return,
while the critic seeks to approximate the value $ \fce \stval s $ of each state $ s $ encountered when following the policy $\pi$.
The critic's prediction is used to guide the actor when training.

% jak se učí PPO
We will denote the critic by $ \critic $\david{tyjo teď se mi vůbec nelíbí to, že se to bude plést s tou policy. Co třeba $\hat{v}(\cdot|\phi)$?}\filip{hele rozleželo se mi to, a zase bych to nechal takhle... přijde mi že je to relativně jasný}\david{jako relativně, ale zkus to schválně dát přečíst třeba Julii. My už víme o co jde a bavíme se o tom dlouho.}\filip{Dal jsem jí to, počkáme co odpoví}, where $ \phi $ represents the parameters of the neural network, and the actor as $ \actor $, again with parameters $ \theta $.
Before defining their respective losses, let us mention how the training process works on a high level.
The critic and the actor are trained simultaneously.
The training process is split into \emph{iterations}.
In each iteration, the agent collects multiple \emph{trajectories} (sequences of state-action-reward triplets, always starting from some initial state\filip{initial state jsme mimochodem nedefinovali....}\david{za mě super, initial state bych buď nedefonoval, nebo dal pod definici returnu za E[G0]?}\filip{Straka má distribuci nad initial stavama... to by asi taky bylo dobrý, a možná to dát už do definice MDP?}) by interaction with the environment, each time using the actor's current strategy, possibly in parallel.
Then, for each time step in each trajectory, we define the so-called \emph{advantage} of the current strategy as $ \hat {A_t} \deq G_t - \fce\critic {s_t} $.
An approximation of $ G_t $ can be easily computed from the observed rewards.
The advantage quantifies how much more reward the agent received from a given state when he chose the action $ a_t $, compared to the prediction of the critic (i.e. approximately the reward he expected under the current strategy $ \actor $)?.
If the advantage is positive, it suggests the agent should play $ a_t $ \emph{more often}, as it lead him to more reward then expected, and if it is negative, it is probably best to play it \emph{less often}.

From the gathered values, we perform $ K $ training epochs on both the actor and the critic.
After the training, we drop all the gathered rewards, and move to the next iteration.

% TODO: proto chceme ty neuronky

% critic dělá tohle, minimalizuje tohle
% TODO: return by měl mít nějaký index říkající z jaké policy vyšel
Let us now discuss the details of the actor and the critic.
The critic is trained to minimize the following loss \[
	\losscritic \deq \fceb {\EE_{\actor,t}}{\left( \return_t - \fce{\critic}{s_t} \right)^2},
\]
where $ \EE_{\pi,t} $ denotes the empirical expectation over all the timesteps $ t $, and over all trajectories in the current iteration.
This loss results in the critic estimating the value of the state $ s_t $ as close as possible.

% actor dělá tohle, minimalizuje tohle,
The actor is trying to learn the policy which maximizes the expected return.
As mentioned above, when $ \hat{A_t} $ is positive, we want to increase the probability with which we choose the action $ a_t $, and decrease it otherwise.
Observe that one way to express this is as maximizing (over the parameters $ \theta $) the following objective \[
	\frac{\fce{\actor}{a_t \suchthat s_t}}{\fce{\actor[\theta_{\text{old}}]}{a_t \suchthat s_t}} \hat{A_t},
\]
where $ \theta_{\text{old}} $ denotes the parameters of the actor from the previous iteration.

However, if we were to maximize this objective directly, with multiple gradient-descent updates, the policy would change too much, and our approximations of $ G_t $, which were gathered using the old policy, would become too inaccurate.
To prevent this, we want to \emph{clip} the loss, so it cannot become \emph{too low} in one iteration.
Hence, the objective of the actor is defined followingly \begin{align*}
	\lossactor &\deq \fceb {\EE_{\actor[\theta_{\text{old}}],t}}{\fcec\min{\left[ \frac{\fce\actor{a_t \suchthat s_t}}{\fce{\actor[\theta_{\text{old}}]}{a_t \suchthat s_t}} \right]^{1+\varepsilon}_{1-\varepsilon} \hat{A_t}, \frac{\fce\actor{a_t \suchthat s_t}}{\fce{\actor[\theta_{\text{old}}]}{a_t \suchthat s_t}} \hat{A_t}}}, \\
	\text{where }
	\left[ x \right]^a_b &\deq \begin{cases}
		a & \text{if $ x > a $,} \\
		b & \text{if $ x < b $,} \\
		x & \text{otherwise.}
	\end{cases}
\end{align*}
\[
\]
The hyperparameter $ \varepsilon $ is used to control the amount of clipping, and thus the maximum rate of change of the policy in each iteration, along with the stability of the learning process.
% In CITACE, the authors suggest the choice of $ \varepsilon = 0.2 $.\david{dal bych spíš jako footnote, jinak to je takový výstřel do tmy.}\filip{a nebo zahodit? :P} % TODO: citace
By taking the minimum of the clipped and unclipped objective, we ensure that the loss is clipped from only one direction.
This is crucial as, if we are unlucky, one policy update might actually \emph{decrease} the objective, and if we were to clip from both sides, we would be unable to learn anything for the remainder of the epochs, since the gradient of the loss in the clipped region is zero.

% PPO pseudokod
\begin{algor}{PPO, Actor-Critic Style}{Clip range $ \varepsilon \in \left( 0,1 \right) $, number of epochs $ K $, number of iterations $ I $, number of trajectories $ R $, $ \alpha_\theta $, $\alpha_\phi $, the coefficients of losses of the actor, critic.}
	{Trained PPO model -- the critic $ \critic $ and actor $ \actor $.}
	\State Initialize $ \critic $ and $ \actor $ randomly.
	\Indent For $ i \in \left\{ 1,2,\ldots,I \right\} $:
		\State Gather rewards for $ R $ trajectories using the current policy $ \actor $.
		\State Compute the advantages $ \hat{A_t} $ for each timestep $ t $ in each trajectory.
		\State Compute $ K $ epochs gradient descent of both the actor and the critic, minimizing the following loss: \[
			\fce{L}{\theta, \phi} \deq \alpha_\theta \lossactor + \alpha_\phi \losscritic.
		\]
	\EndIndent
\end{algor}

% PPO vs ostatní AC -- problém když mezi iteracema se hodně změní policy -> kritik je kokot -> nestabilní
Finally, let us mention the differences between PPO and other RL algorithms.
We have already mentioned that we require an algorithm which works well with continuous state spaces, however, this is not specific to PPO, but rather to the fact that we are using neural networks for the policy. \filip{tohle ocitovat?}\david{jako citovat, že neuronky berou na vstupu reálná čísla?}\filip{:D spíš jsem myslel odkazy na ostatní přístupy které fungují s continuous stavy, třeba REINFORCE}
PPO operates on an \emph{on-policy} basis, meaning thath the critic always learns from data generated by the current policy of the actor.
This is in contrast to \emph{off-policy} algorithms, which can utilize rewards gathered in previous iterations under previous policies to train the new policy.
On-policy algorithms generally maintain better learning stability since the critic isn't influenced by outdated rewards.
Conversly, off-policy algorithms are more data efficient, as utilizing old trajectories reduces the need to gather as many new ones\filip{s touhle větou ještě nejsem spokojenej}.\filip{citace}\david{něco zkusim najít}
PPO remedies the inefficiency of throwing away old trajectories by parallelizing the data gathering process, and thus allowing much faster gathering of new trajectories.
The gathering of trajectories is typically performed on CPUs, which are readily available on modern computation centers.
\david{Něco o tom, že sbírání trajektorií se typicky dělá na cpu a to je levný? Nebo prostě to, že ten algoritmus jde obecně škálovat?}
\filip{V jakém směru škálovat? :D}
\filip{Uuuuuh, myslíš tím že proč je ta $ \lossactor $ tak "divná" (oproti tomu logaritmu jako je u REINFORCE s baseline)? Nebo zmínit vůbec to že to vlastně jde i bez kritika, jako nějakej $ \varepsilon $-greedy?}
These properties make PPO a good choice for various problems, including the one we are interested in, as described in \Cref{sec:ppo}.
\filip{Jaký to je takhle?}



% Finally, let us mention the differences between PPO and other RL algorithms.
% We have already mentioned that we require an algorithm which works well with continuous state spaces, however, this is not specific to PPO, but rather to the fact that we are using neural networks for the policy. \filip{tohle ocitovat?}\david{jako citovat, že neuronky berou na vstupu reálná čísla?}\filip{:D spíš jsem myslel odkazy na ostatní přístupy které fungují s continuous stavy, třeba REINFORCE}
% PPO is so-called \emph{on-policy}, as the critic always learns from data generated by the current actor's policy.
% This is in contrast to \emph{off-policy} algorithms, which can use rewards gathered by an old policy to train a new one.
% On-policy algorithms generally have better learning stability, as the critic isn't \textquote{mislead} by no longer relevant rewards.
% On the other hand, off-policy algorithms are generally more data efficient, since they don't throw away old trajectories completely, and hence don't need to gather as many new ones.\filip{citace}\david{něco zkusim najít}
% PPO remedies the inefficiency of throwing away old trajectories by parallelizing the data gathering process, and thus allowing much faster gathering of new trajectories.
% \filip{potřebuju to nějak zakončit ale tohle je strašný:}\david{Něco o tom, že sbírání trajektorií se typicky dělá na cpu a to je levný? Nebo prostě to, že ten algoritmus jde obecně škálovat?}
% \filip{V jakém směru škálovat? :D}
% In conclusion, PPO is designed for stability of learning, while offering parallelization of data gathering, which makes it also fast, as it is limited only by the number of CPUs, which tend to be readily available on current computation centers.
% \david{tady by se fakt hodilo mluvit o tom policy gradient vs jak trénuje actora ppo.}
% \filip{Uuuuuh, myslíš tím že proč je ta $ \lossactor $ tak "divná" (oproti tomu logaritmu jako je u REINFORCE s baseline)? Nebo zmínit vůbec to že to vlastně jde i bez kritika, jako nějakej $ \varepsilon $-greedy?}

% This approach , and it generally leads to more stable learning.\filip{as the critic isnt confused by the old actor's choices? něco takového}\david{A pak bych spíš napsal něco jako "when dealing with off-policy trajectories, one needs to resort to importance sampling, which introduces variance (citace imp samplingu)"}
% Another interesting property we have briefly mentioned already is the \emph{on-policy} approach of PPO.
% This means that after each update of the policy (each iteration of the PPO algorithm), we drop all the gathered rewards, and start anew.
% \emph{Off-policy} algorithms work well when the expected returns do not change much between iterations, however, when they do, the critic's predictions become inaccurate, and the training becomes unstable.\david{On the other hand, off-policy algorithms can be much more data-efficient, as they are able to reuse old trajectories (possibly generated by some expert which is inaccessable at train-time).}
% PPO further stabilizes the training by clipping the loss, as discussed above.
% \filip{Tenhle paragraf je takovej meh, a chtěl by asi nějak pěkně ukončit, ale nevím jak}
% \david{tohle kdyžtak nějak rozepíšeme. PPO je rozhodně ještě super díky tomu, že jde paralelizovat to sbírání trajektorií. To vlastně dost zachranuje to, že jinak je off-policy daleko víc data efficient.}
% We will discuss the differences between on-policy and off-policy algorithms later.

\section{Regret Minimization}
\filip{tohle přepsat nebo smazat idk}

% cil RM
Regret minimization is a technique used in online learning, where the agent tries to pick a strategy, that would have been good in hindsight.
By hindsight, we mean that the agent compares itself other sequences of strategies he could have chosen, and tries to minimize the difference between the losses he experienced, and the losses the best sequence of strategies would have experienced.

The formal setup of regret minimization differs in a few aspects from the general online learning setup given in \Cref{def:mdp}.
Let us first define the setup for a single-state environment, and then we will generalize it to the multi-state case.

\begin{defi}[Regret minimization environment]
	A \emph{regret minimization environment} is a tuple $ \left( \actions, \ls \right) $, where \begin{enumerate}[ ]
		\item $ \actions $ is a non-empty set of \emph{actions},
		\item $ \vec \ls: \dist \actions \to \R^n $ is the \emph{loss function}.
	\end{enumerate}
\end{defi}

Notice that unlike \Cref{def:mdp}, where we had to choose one action at each step,
here we choose a distribution over the actions, and receive a per-action loss, based on that distribution.
This gives us extra information about \textquote{which action costs us the most}.
Note however that this doesn't mean that the action with lowest loss is the best action to choose.
Take for example the case of rock, paper, scissors.
If we supply a uniform distribution, we will receive a per-action loss of $ \frac{1}{3} $ for each action.
However, if we only played scissors, for example, we would suddenly always receive a loss of $ 1 $ on it!
\filip{Nejsem si jistej jestli ten example vyzněl tak jak měl xD}

% experienced loss
To quantify the loss received by the agent, we will compute the expected value of the per-action loss, given the distribution chosen by the agent.

\begin{defi}[Experienced loss]
	Let $ \vec{\ls} $ be a per-action loss received by the agent when playing $ \pi \in \dist \actions $.
	The \emph{experienced loss} of the agent is \[
		\vec\xl \deq \E_{\pi} \left[ \sum_{i=1}^{n} \pi_i \ls_i \right].
	\]
\end{defi}

The new definition of the loss function also leads to a new definition of the cumulative loss.

\begin{defi}[Cumulative loss, regret]
	Suppose the agent received losses $ \vls^1, \dots, \vls^T $.
	Then the \emph{cumulative loss} of the agent is \[
		\vcxl^T \deq \sum_{t=1}^{T} \vls^t.
	\]
\end{defi}


We would now like to compare our agent against some \emph{comparison set} $ \algs $ of online algorithms.
An online algorithm can be thought of as a function $ \alg $, which maps a sequence of already received losses to a distribution over the actions.
To properly quantify this, we will define so-called \emph{regret} for a general comparison set $ \algs $, and later discuss behavior of regret for specific comparison sets.

% regret
\begin{defi}[Regret]
	Let $ \algs $ be a comparison set.
	Let $ \vls^1, \dots, \vls^T $ be a sequence of losses received by the agent, playing a sequence of strategies $ \pi^1, \dots, \pi^T $.
	Then the \emph{regret} of the agent after $ T $ steps, with respect to $ \algs $, is \[
		R^T_{\algs} \deq \cxl^T  - \min_{\alg \in \algs} \cxl^T_{\alg},
	\]
	where $ \cxl^T_{\alg} $ is the cumulative experienced loss of the algorithm $ \alg $ computed from the losses $ \vls^i $ received by the agent.
\end{defi}

% předpokládáme že losses zůstanou stejné
% TODO: někde popsat že vektory jsou tučně
Let us emphesize that experienced losses of the algorithms from $ \algs $ are computed from $ \vls^1, \dots, \vls^T $.
This means, that those aren't the losses the algorithms would have actually experienced, had they been played instead of the agent, as the loss $ \vls^i $ depended on the the distribution $ \pi^i $ chosen by the agent.
Consider the simple adversarial example, where the loss $ \ls_i $ is $ 1 $ for all actions with non-zero probability, and $ 0 $ otherwise.
Suppose that the agent chooses action $ 1 $ with probability $ 1 $, and an algorithm $ \alg $ chooses action $ 2 $ instead.
Then the experienced loss of the agent is $ 1 $, while the experienced loss of $ \alg $, as seen from the definition of regret, is $ 0 $, even though, had we followed $ \alg $, we would have experienced loss $ 1 $, too.
% TODO: upřsenit tenhle příklad
% \david{Tomu konci nerozumím. Loss akce 2 je 0 ne? Takže regret k té akci je 1.}
% \filip{Přesně tak. Já tam akorát mluvím o loss. Chci tam vysvětlit že když počítám $\cxl^T_{\alg} $, tak to není to samý jako co by dostal ten algoritmus kdyby reálně hrál.}
% \david{Aha, už chápu. To je určitě fajn point vypíchnout, ale trochu se to ztrácí v téhle formulaci.}

We won't want to compare against all possible algorithms, because among them, there will be one that always guesses the best action, which will lead to the lowest possible loss.
For an adversarial environment, this is impossible to compare against.
% TODO: přeformulovat
% \david{Tohle není pravda - porovnat se můžeš, akorát takovýhle regret nejde minimilizovat proti libovolnému black-boxu. Tohle diskuzi bych odložil po definici regretu.}

% swap regret, internal, external?
Let us now mention different types of regret, meaning different comparison sets $ \algs $, along with their properties.
The first and probably most studied type of regret is the \emph{external regret}.
Many times in literature, when the term \emph{regret} is used without any further specification, it is meant to be the external regret.\david{Přesně z tohohle důvodu bych používal instantaneous / cumulative regret. Jen regret je možná moc vágní, kor takhle na začátku v refinicích.}
It is defined on the comparison set $ \algs_\actions \deq \left\{ \alg_a \suchthat a \in \actions \right\} $, where $ \alg_a $ is the algorithm that always chooses action $ a $.
The external regret analyzes what would have happened, had the agent chosen a single action, and stuck with it for the entire game.\david{Tady game není jasné co vůbec je. Možná bych tomu online procesu říkal nějak úplně jinak, ale nenapadá mě dobré jméno.}
The formula can be simplified to \[
		R^T_{\algs_\actions} = \cxl^T - \min_{a \in \actions} \cl^T_a.
	\]

For stronger comparison sets, we will use algorithms which are derived from the strategy of the agent, but modify it using \emph{modification rules}.
A modification rule is any function $ F: \actions \to \actions $, and the resulting algorithm $ \alg_F $ looks at the strategy $ \pi $ of the agent, and plays $ \pi' $, defined as \[
	\pi'_a \deq \sum_{i: F(i) = a} \pi_i.
\]
Consider the example of a modification rule $ F_{i,j} $, which respects the strategy of the agent, but always plays action $ j $ instead of action $ i $.
Formally: \[
	\fce{F_{i,j}}{k} = \begin{cases}
		j & \text{if } k = i, \\
		k & \text{otherwise.}
	\end{cases}
\]
% TODO: víc popsat?
The regret with respect to the comparison set $ \left\{ \alg_{F_{i,j}} \suchthat i,j \in \actions \right\} $ is called the \emph{internal regret}.
Even more generally, if we consider a comparison set with all possible modification rules, we get the \emph{swap regret}.
% TODO: vlastnosti regretů?

% regret minimizer
Now that we have a way to measure how well the agent is doing, we need to find a way to adapt the strategy of the agent, so that the regret is minimized.
Ideally, we would like to find a strategy, which keeps regret $ 0 $, but as we have already seen with the adversarial example, that is in general impossible.
Instead, we will try to find a strategy, which keeps the regret as small as possible.
More specifically, we are aiming for a strategy, which keeps the regret sublinear in $ T $, that is, $ R^T_{\algs} = \fce{o}{T} $.

\begin{defi}[ ]
	Let $ \algs $ be a comparison set.
	An agent is an \emph{$ \algs $-regret minimizer}, if its $ \algs $-regret satisfies \[
		R^T_\algs = \fce{o}{T}.
	\]
\end{defi}
Once again, when talking about a regret minimizer without any further specification, we mean an external regret minimizer.
External regret is special, because there are algorithms that can provably achieve sublinear regret, and we will discuss them in further sections.
% TODO: Balko thm 2.55.
% TODO: proč sub-lineární regret je cool

% TODO: příklady regret minimizerů?

% TODO: sequence form game

% TODO: predictive regret?

% TODO: příklad (asi ten s deštníkama?)
We can imagine that the function comes from a distribution of functions.
This distribution is unknown to the agent, though gathers information about it from the losses he receives as he plays the game.


% TODO: minimalizovat regret a maximalizovat reward je to stejné proti "oblivious environment" (to je to, kde bereš gradient loss podle policy jako 0)
% TODO: RM sublin regret, ale ne minimalizace:
